Accelerating parameter inference with graphics processing units

### 2019-01-30 paper/paper.tex
paper.tex work on BNS section

\title{Accelerating parameter inference with graphics processing units}
\author{D. Wysocki}
\affiliation{\RIT}
\author{Yao-Lung L. Fang}
%\author{Yao-Lung L. Fang (方耀龍)}
\affiliation{Computational Science Initiative, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affiliation{National Synchrotron Light Source II, Brookhaven National Laboratory, Upton, NY 11973, USA}
\author{R. O'Shaughnessy}
\affiliation{\RIT}
\author{Jake - if he does runs}
\affiliation{\RIT}
\begin{abstract}
Gravitational wave Bayesian parameter inference involves repeated comparisons of GW data to generic candidate predictions.
Even with algorithmically efficient methods like RIFT or reduced-order quadrature, the time needed to perform these
calculations and overall computational cost can be significant compared to the minutes to hours needed to achieve the
goals of low-latency multimessenger astronomy.  By translating some elements of the RIFT algorithm to operate on
graphics processing units (GPU), we demonstrate substantial performance improvements, enabling dramatically reduced
overall cost and latency. 
\end{abstract}
\maketitle

@@

\subsection{Exploiting low-cost likelihoods}
Previously, each instance of ILE examined one intrinsic point $\lambda$.  The overall cost of this ILE evaluation
involved two parts: a startup cost, a setup cost, and a Monte Carlo integration cost.  The  startup cost $\tau_{start}$
is associated with code setup followed by reading, data conditioning, and Fourier-transforming the
data.  The setup cost $\tau_{setup}$ arises from waveform generation and  the inner products $U,V,Q$.  Finally, the Monte Carlo
cost $\tau_{mc}=N_{ad} \tau_{ad}+N_{it}\tau_{eval}$ increases with the number of Monte Carlo iterations $N_{it}$ with
and the number of times $N_{ad}$ the sampling prior used in adaptive Monte Carlo is regenerated, in proportion to their
cost.   In the standard configuration, the adaptive sampling prior is regenerated every $n_{\rm chunk}$ iterations (i.e,
 $N_{ad} \simeq N_{it}/n_{chunk}$), regenerating a sampling prior in two sky location parameters and distance.  
The choice of one $\lambda$ for each instance of ILE was eue to the substantial time  $\tau_{MC}$, which vastly
dominated the overall computational cost.   For example, for a typical analysis of a short signal --  a typical binary
black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$ --  this version has
cost elements as shown in Table \ref{tab:CostBreakdown}, based on the assumption that the MC terminates after $N_{\rm it}\simeq 2\times 10^6$ iterations using
$n_{\rm chunk}\simeq 10^4$.  
The marginalized likelihood described above  converges incredibly rapidly to a small value if the candidate model is
inconsistent with a signal (or the absence thereof) in the data, with only $N_{it} \simeq O(10^4)$  evaluations needed.

\begin{table}
\begin{tabular}{lrr|ccccc|r}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ & $\frac{T_{ILE}}{N_{eval}}$\\  %\hline 
  &  & Hz & sec & sec & & $\mu$sec & $\mu$sec\\ \hline 
CPU & 16384 & $(2,\pm 2) $ & 20 & && & &  \editremark{900}  \\ 
       & 4096 & $(2,\pm 2) $ &   20 \\ \hline
% LIGO-WA: 

@@

GPU (a) & 16384 & $(2,\pm 2) $  & 20  & &   \\
            & 4096 &$(2,\pm 2) $  &  20 &\\

@@

GPU (b) & 16384 & & 20  & 1.8 & 1 & 0.85& 12 &28\\
       & 4096 & & 20 & $1.2 $ &  1  & 0.75 &  & 25\\ \hline
GPU (b) & 16384 & $\pm 2, \pm 1$ &  \\
       & 4096 & $\pm 2, \pm 1$ \\

@@

\caption{\label{tab:CostBreakdown}Evaluation costs on default hardware, for a two-mode system $(l,m)=(2,\pm 2)$.  Scenario (a) is a compute node on LIGO-WA; scenario (b) is more modern hardware. }
\end{table}

With the new low-cost likelihood evaluations, however, the startup and setup costs  $\tau_{start},\tau_{setup}$ now can be
comparable or in excess of the total time used to evaluate the likelihood.   In fact, the total time per Monte Carlo
evaluation $\tau_{it}=\tau_{manage}+\tau_{eval}$  spent in general-purpose overhead $\tau_{manage}$ will be much larger than the time $\tau_{eval}$ spent
carrying out scientific calculations by evaluationg the likelihood.  For these reasons, we reorganize the workflow, so each
instance of ILE loops over
$N_{eval}$ different choices for $\lambda$. 
%
Additionally, particularly for massive binary black holes, we investigate two additional performance improvements.  First and
foremost, we lower the sampling rate, thus lowering the startup cost $\tau_{start}$ and particularly setup cost
$\tau_{setup}$.  Previously and out of an abundance of caution, \ILE{} employed a sampling rate $1/\Delta t = 16384
\unit{Hz}$ for all calculations, but terminated all inner products in Eq. (\ref{eq:overlap}) at roughly $f_{\rm
  max}=2048\unit{Hz}$ or less.   Reducing the sampling rate by a factor $s$ will the  cost of all operations with
timeseries -- they are shorter.  Depending on the relative cost of overhead versus array operations,  $\tau_{setup}$
\emph{and} the cost per evaluation $\tau_{eval}$ decrease modestly because of this effect.  
Also, to insure safe inner products over short data sets in the presence of very narrow spectral lines, we operated on a
significantly-padded data buffer,
using 
``inverse spectrum truncation'' \editremark{citation}.  For binary black holes, the degree of padding was significantly
in excess of the signal duration \editremark{numbers: say 64s for a BBH!}.  
Instead and following LI, we adopt a much shorter inverse spectrum truncation length, a much shorter padding buffer, and
a Tukey window applied to the data to remove discontinuities at the start and end of each segment.   
Reducing the duration of data analyzed by a factor $s'$ will reduce the cost $\tau_{setup}$ by a factor $s'$.  
%
Combining these factors together, the overall computational cost of each marginalized likelihood
evaluation $T_{ILE}/N_{eval}$  is
\begin{align}
T_{ILE}/N_{\rm eval} &= \frac{\tau_{start}}{N_{\rm eval}} 
 + 
 [ \tau_{setup}/s' + N_{ad} \tau_{ad} + N_{it}\tau_{it}
 ] 
\end{align}
For a typical binary black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$, this
version requires $\tau_{start} \simeq 12\unit{s}$, $\tau_{setup}/ss' \simeq XX$, $\tau_{ad}\simeq XXX$, and $N_{it}\tau_{\it} \simeq XX$
\editremark{numbers}.  
Because of the extremely high cost of adaptive overhead, we only adapt in two parameters, corresponding to sky location; 
and only adapt during the first iteration which returns significant $\ln {\cal L}_{\rm marg}$.  As a result, $N_{ad}$
(the number of adaptive stages \emph{per likelihood evaluation}) scales as $N_{it}/N_{chunk}/N_{\rm eval}$ and does not
increase with the overall number of samples, so the overall evaluation time scales as
\begin{align}
T_{ILE,mod}/N_{\rm eval} &= \frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \nonumber \\ &
 + 
 [ \tau_{setup}/s'  + N_{it}\tau_{it}
 ] 
\end{align}
Typically we target $N_{\rm eval} \simeq 20$ so that startup and other one-time costs  are not a significant contributor to the overall
runtime.   In this configuration, the run time per marginalized likelihood is around $T_{ILE}/N_{\rm eval} \simeq 25
\unit{s}$ for a binary black hole \editremark{checkme - old number}.    Of this $25 \unit{s}$, only roughly
$1.5\unit{s}$ corresponds to evaluating the likleihood, implying the cost $\tau_{it} \simeq 1.5/N_{\rm it} \simeq
7.5\times 10^{-7}$.
Likewise, because overhead dominates over array manipulations, an analysis with  $16\unit{kHz}$ timeseries requires only
marginally more time than the corresponding $4\unit{kHz}$ analysis.

For low-latency analyses, we need to assess the overall wallclock time needed to perform an analysis, often with a
restricted number $N_{GPU}$ of available GPU-enabled machines.   To  complete $N_{\rm net}$ likelihood evaluations
with these resources requires
\begin{align}
T_{net,mod} &= \frac{N_{\rm net}}{N_{\rm GPU} N_{\rm eval}} T_{ILE,mod} \nonumber \\
& =
  \frac{N_{net}}{N_{GPU}}\left[
\frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \right. \nonumber \\
& \left. + 
 [ \tau_{setup}/ss'  + N_{it}\tau_{it}/s
 ] 
 \right]
\end{align}
Typically we target $N_{\rm eval}  \lesssim 3\times 10^4, N_{\rm GPU} \simeq 100$, or roughly $300\times$ the cost per
individual marginalized likelihood evaluation.  For the binary black hole configuration described above, marginalized
likelihood evaluations will complete in about $100 \unit{minutes}$ \editremark{checkme}. This number can be reduced to
tens of minutes with a modestly larger GPU pool, or by marginally more conservative convergence thresholds..
This discussion ignores the latency introduced by the  gaussian process interpolation stage at the end.  We will revisit
accelerated GP interpolation in subsequent work.

@@

\editremark{Highly heterogeneous GPU environment - very different performance depending on available resources}

\begin{figure}

\end{figure}

* Comparison: agreement.  Sky localization example

* Benchmarking performance (new speed for BBHs; also show BNS with tides)

>>>>

\title{Accelerating parameter inference with graphics processing units}
\author{D. Wysocki}
\affiliation{\RIT}
\author{Yao-Lung L. Fang}
%\author{Yao-Lung L. Fang (方耀龍)}
\affiliation{Computational Science Initiative, Brookhaven National Laboratory, Upton, NY 11973, USA}
\affiliation{National Synchrotron Light Source II, Brookhaven National Laboratory, Upton, NY 11973, USA}
\author{R. O'Shaughnessy}
\affiliation{\RIT}
\author{Jake?}
\affiliation{\RIT}
\begin{abstract}
Gravitational wave Bayesian parameter inference involves repeated comparisons of GW data to generic candidate predictions.
Even with algorithmically efficient methods like RIFT or reduced-order quadrature, the time needed to perform these
calculations and overall computational cost can be significant compared to the minutes to hours needed to achieve the
goals of low-latency multimessenger astronomy.  By translating some elements of the RIFT algorithm to operate on
graphics processing units (GPU), we demonstrate substantial performance improvements, enabling dramatically reduced
overall cost and latency. 
\end{abstract}
\maketitle

@@

We illustrate the code using two synthetic signals: a binary black hole with masses $m_{1,2}=35,30 M_\odot$ and a binary
neutron star with masses $1.4 M_\odot$ and $1.35 M_\odot$.
We generate  synthetic data for a two-detector LIGO configuration, assuming operation at  initial (BBH) or advanced (BNS) design  sensitivity.
Both signals are chosen to have peak SNR $\simeq 14$ and are generated with SEOBNRv4.    We perform a multi-stage iterative RIFT analysis of these signal with the
time-domain SEOBNRv4 (BBH) or TaylorT4 (BNS) approximation, under the simplifying assumption that both objects are point particles with zero
spin.  We repeat this analysis on heterogeneous LVC collaboration computing resources, to assess our variable
performance across architectures.  Notably, we investigate the following GPU options: (a) GTX 1050 Ti at LIGO-CIT,
available in large numbers; (b) V100, available on selected high-performance machines; (c) and
GTX 1050 at LIGO-LHO.    Unless otherwise noted in the text, we will discuss code configurations using CPU-only and GPU (b) using
$4096\unit{Hz}$ sampling and only the  $\ell=2, m=\pm 2$ modes.

\subsection{Massive compact binary black holes}
Previously, each instance of ILE examined one intrinsic point $\lambda$.  The overall cost of this ILE evaluation
involved two parts: a startup cost, a setup cost, and a Monte Carlo integration cost.  The  startup cost $\tau_{start}$
is associated with code setup followed by reading, data conditioning, and Fourier-transforming the
data.  The setup cost $\tau_{setup}$ arises from waveform generation and  the inner products $U,V,Q$.  Finally, the Monte Carlo
cost $\tau_{mc}=N_{ad} \tau_{ad}+N_{it}\tau_{eval}$ increases with the number of Monte Carlo iterations $N_{it}$ with
and the number of times $N_{ad}$ the sampling prior used in adaptive Monte Carlo is regenerated, in proportion to their
cost.   In the standard configuration, the adaptive sampling prior is regenerated every $n_{\rm chunk}$ iterations (i.e,
 $N_{ad} \simeq N_{it}/n_{chunk}$), regenerating a sampling prior in two sky location parameters and distance.  
The choice of one $\lambda$ for each instance of ILE was eue to the substantial time  $\tau_{MC}$, which vastly
dominated the overall computational cost.   For example, for a typical analysis of a short signal --  a typical binary
black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$ --  this version has
cost elements as shown in Table \ref{tab:CostBreakdown}, based on the assumption that the MC terminates after $N_{\rm it}\simeq 2\times 10^6$ iterations using
$n_{\rm chunk}\simeq 10^4$.  
The marginalized likelihood described above  converges incredibly rapidly to a small value if the candidate model is
inconsistent with a signal (or the absence thereof) in the data, with only $N_{it} \simeq O(10^4)$  evaluations needed.

\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% ~/parse_report.sh profile_nogpu_pcdev13.log | more
CPU & 16384 & $\pm 2 $ & 20 & &&532 & &  700  \\ 
       & 4096 & $\pm 2 $ &   20 \\ \hline
% ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
%    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
%   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & &  1000  \\ 
       & 4096 & $ \pm 2, \pm 1 $ &   20 \\ \hline
% LIGO-WA, or a compute node at CIT with a grpahics catd: 

@@

%  - CIT jobs, on compute nodes
%        20 points at 16kHz take about 90 min
%        20 points at kHz take anywhere from
%        condor_userlog iteration_0_ile/logs/*.log
GPU (a) & 16384 & $(2,\pm 2) $  & 20  & & && & 270 \\
            & 4096 &$\pm 2 $  &  20 &  & & && 45 \\ \hline

@@

GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 12 &28 & 15\\
       & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 &  & 25\\ \hline
% ~/parse_report.sh profile_gpu_HM_lowres_pcdev13.log | more
GPU (b) & 16384 & $\pm 2, \pm 1$ & 20  \\
       & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  && 35 & \\ \hline
% LHO or LLO worker nodes
%  ./parse_report.sh 20190130-profile_gpu_LHOworker_16kHz.log | more
GPU (c) & 16384 & $\pm 2 $  & 20  & & & 18& &160 &  \\
            & 4096 &$\pm 2 $  &  20 &  & & 22  & & 140 & $\simeq 50$ \\

@@

\caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
  marginalized likelihood on default
  hardware, for a two-mode system $(l,m)=\pm 2$ analysing $T=8\unit{s}$ of data with a massive binary black hole
  $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.}
\end{table*}

With the new low-cost likelihood evaluations, however, the startup and setup costs  $\tau_{start},\tau_{setup}$ now can be
comparable or in excess of the total time used to evaluate the likelihood.   In fact, the total time per Monte Carlo
evaluation $\tau_{it}=\tau_{manage}+\tau_{eval}$  spent in general-purpose overhead $\tau_{manage}$ will be much larger than the time $\tau_{eval}$ spent
carrying out scientific calculations by evaluationg the likelihood.  For these reasons, we reorganize the workflow, so each
instance of ILE loops over
$N_{eval}$ different choices for $\lambda$. 
%
Additionally, particularly for massive binary black holes, we investigate two additional performance improvements.  First and
foremost, we lower the sampling rate, thus lowering the startup cost $\tau_{start}$ and particularly setup cost
$\tau_{setup}$.  Previously and out of an abundance of caution, \ILE{} employed a sampling rate $1/\Delta t = 16384
\unit{Hz}$ for all calculations, but terminated all inner products in Eq. (\ref{eq:overlap}) at roughly $f_{\rm
  max}=2048\unit{Hz}$ or less.   Reducing the sampling rate by a factor $s$ will the  cost of all operations with
timeseries -- they are shorter.  Depending on the relative cost of overhead versus array operations,  $\tau_{setup}$
\emph{and} the cost per evaluation $\tau_{eval}$ decrease modestly because of this effect.  
Also, to insure safe inner products over short data sets in the presence of very narrow spectral lines, we operated on a
significantly-padded data buffer,
using 
``inverse spectrum truncation'' \editremark{citation}.  For binary black holes, the degree of padding was significantly
in excess of the signal duration \editremark{numbers: say 64s for a BBH!}.  
Instead and following LI, we adopt a much shorter inverse spectrum truncation length, a much shorter padding buffer, and
a Tukey window applied to the data to remove discontinuities at the start and end of each segment.   
Reducing the duration of data analyzed by a factor $s'$ will reduce the cost $\tau_{setup}$ by a factor $s'$.  
%
Combining these factors together, the overall computational cost of each marginalized likelihood
evaluation $T_{ILE}/N_{eval}$  is
\begin{align}
T_{ILE}/N_{\rm eval} &= \frac{\tau_{start}}{N_{\rm eval}} 
 + 
 [ \tau_{setup}/s' + N_{ad} \tau_{ad} + N_{it}\tau_{it}
 ] 
\end{align}
For a typical binary black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$, this
version requires $\tau_{start} \simeq 12\unit{s}$, $\tau_{setup}/ss' \simeq XX$, $\tau_{ad}\simeq XXX$, and $N_{it}\tau_{\it} \simeq XX$
\editremark{numbers}.  
Because of the extremely high cost of adaptive overhead, we only adapt in two parameters, corresponding to sky location; 
and only adapt during the first iteration which returns significant $\ln {\cal L}_{\rm marg}$.  As a result, $N_{ad}$
(the number of adaptive stages \emph{per likelihood evaluation}) scales as $N_{it}/N_{chunk}/N_{\rm eval}$ and does not
increase with the overall number of samples, so the overall evaluation time scales as
\begin{align}
T_{ILE,mod}/N_{\rm eval} &= \frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \nonumber \\ &
 + 
 [ \tau_{setup}/s'  + N_{it}\tau_{it}
 ] 
\end{align}
Typically we target $N_{\rm eval} \simeq 20$ so that startup and other one-time costs  are not a significant contributor to the overall
runtime.   In this configuration, the run time per marginalized likelihood is around $T_{ILE}/N_{\rm eval} \simeq 25
\unit{s}$ for a binary black hole.    Of this $25 \unit{s}$, only roughly
$1.5\unit{s}$ corresponds to evaluating the likleihood, implying the cost $\tau_{it} \simeq 1.5/N_{\rm it} \simeq
7.5\times 10^{-7}$.
Likewise, because overhead dominates over array manipulations, an analysis with  $16\unit{kHz}$ timeseries requires only
marginally more time than the corresponding $4\unit{kHz}$ analysis.

For low-latency analyses, we need to assess the overall wallclock time needed to perform an analysis, often with a
restricted number $N_{GPU}$ of available GPU-enabled machines.   To  complete $N_{\rm net}$ likelihood evaluations
with these resources requires
\begin{align}
T_{net,mod} &= \frac{N_{\rm net}}{N_{\rm GPU} N_{\rm eval}} T_{ILE,mod} \nonumber \\
& =
  \frac{N_{net}}{N_{GPU}}\left[
\frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \right. \nonumber \\
& \left. + 
 [ \tau_{setup}/ss'  + N_{it}\tau_{it}/s
 ] 
 \right]
\end{align}
Typically we target $N_{\rm eval}  \lesssim 3\times 10^4, N_{\rm GPU} \simeq 100$, or roughly $300\times$ the cost per
individual marginalized likelihood evaluation.  For the binary black hole configuration described above, marginalized
likelihood evaluations will complete in about $100\; \unit{minutes}$ on a cluster of (b), or more realistically 200
minutes on a cluster of (a). . This number can be reduced to
tens of minutes with a modestly larger GPU pool, or by  marginally more conservative convergence thresholds. on $N_{\rm
  eval}$ or $N_{it}$.  
This discussion ignores the latency introduced by the  gaussian process interpolation stage at the end.  We will revisit
accelerated GP interpolation in subsequent work.

@@

The \ILE{} likelhood involves sums over modes, with the most expensive likelihood and setup operations (involving $Q$) growing linearly with
the number of modes used.  This increased cost is most apparent in Table \ref{tab:CostBreakdown} when comparing the
columns corresponding  $\tau_{it,like}$ for GPU (b) between the $\pm 2$ and $(\pm 2,\pm 1)$ rows.  The likelihood
evaluation cost of runs with 4 modes is roughly twice that of runs with two modes, as expected.

In a heterogeneous computing environment, code performance varies substantially with available resources, as seen by
contrasting corresponding results for (a,b,c) in Table \ref{tab:CostBreakdown}.      For
the slowest hardware,  the overall runtime will be only a factor few smaller than the corresponding CPU-only runtime.

\subsection{Binary neutron stars}
\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
CPU & 16384 & $\pm 2,\pm 1 $ & 20 &  &&  & &    \\ 
       & 4096 & $\pm 2, \pm 1$ & 20 & 1&&   &&  & \\ \hline
GPU (b) & 16384 & $\pm 2, \pm 1$   & 20 & 26 & 2.4 && &35 &  \\
       & 4096 & $ \pm 2, \pm 1 $        &   20  & 11.5 & 2.4 & & &24 & \\ \hline
\end{tabular}
\caption{\label{tab:CostBreakdown:BNS}\textbf{Profiling performance: Binary neutron stars}: Evaluation costs for the
  marginalized likelihood on default
  hardware,  analysing $T=8\unit{s}$ of data with a binary neutron stars
  $m_1=1.4 M_\odot,M_2=1.35 M_\odot$, with convergence threshold $n_{\rm eff} > 50$.
}
\end{table*}

Table \ref{tab:CostBreakdown: BNS} shows the corresponding performance breakdown for our fiducial binary neutron star,
which has signal SNR=32.  In this analysis, we have adopted the TaylorT4 model with $\ell=2$ and all $m=\pm 2,\pm 1$ modes as our
fiducial analysis template.   Despite using more modes and longer signals, and despite an overall likelihood evaluation
cost $\tau_{it}$ consistent BBH analyses,  likelihood evaluations for BNS are often
\emph{as fast or faster} than the corresponding likelihood evaluations for BBH with otherwise comparable settings.

Low-mass compact binaries like binary neutron stars have much longer inspiral times from the same starting frequency of
$20\unit{Hz}$: roughly $160 \unit{s}$.   Nominally, the manipulation of a corresponding power-of-two duration data ($256\unit{s}$) involves
$32\times$ more costly fourier transforms and time integrations than the BBH analysis described above.  
%  a = np.zeros(16384*256,dtype=np.complex64); a.nbytes/1024/1024
[An array of $256\unit{s}$ of data at $16\unit{kHz}$ corresponds to  32 Mb.]  Comparing $\tau_{setup}$ appearing in Tables
 \ref{tab:CostBreakdown} and  \ref{tab:CostBreakdown:BNS}, we indeed find our BNS analysis requires  a factor of order   $32\times$
 longer to set up all necessary inner products.  Unlike the BBH analysis, the inner product evaluation costs in
 $\tau_{setup}$ now dominates the
 overall evaluation time $T_{ILE}/N_{\rm eval}$.

Improved performance arises instead from re-using our adaptive integrator, which (after the first marginalized
likelihood evaluation) can exploit tight localization afforded by the many BNS cycles available in band and the high
amplitude of our fiducial BNS signal.  In other
words, the Monte Carlo integral converges to our target accuracy in of order $10\times $ fewer steps than for BBH.  
Because this improved performance relies on the sampler, not on our choice of hardware, we achieve a similar substantial
improvement for all choices of GPU.

\editremark{test on other hardware}
